Modelling M/EEG data with Bayesian nonparametric multilevel models

### 2025-02-06 brms_meeg.tex
now modelling actual decoding results (instead of simulated ones)

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

Here are some useful references to be discussed
(\citeproc{ref-combrisson_exceeding_2015}{Combrisson \& Jerbi, 2015};
\citeproc{ref-ehinger_unfold_2019}{Ehinger \& Dimigen, 2019};
\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021},
\citeproc{ref-frossard2022}{2022}; \citeproc{ref-gramfort2013}{Gramfort,
2013}; \citeproc{ref-hayasaka_validating_2003}{Hayasaka, 2003};
\citeproc{ref-luck_how_2017}{Luck \& Gaspelin, 2017};
\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
\citeproc{ref-pedersen_hierarchical_2019}{Pedersen et al., 2019};
\citeproc{ref-pernet2015}{Pernet et al., 2015};
\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al., 2023};
\citeproc{ref-rousselet_using_2025}{Rousselet, 2025})\ldots{} See also
(\citeproc{ref-maris2011}{Maris, 2011})\ldots{} and
(\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}) (history of
cluster-based approaches and using a data split?)\ldots{} Cluster
failure (\citeproc{ref-eklund2016}{Eklund et al., 2016})\ldots{}

@@

\[
\begin{aligned} 
y_{i} &\sim \mathrm{EF}\left(\mu_{i}, \phi\right)\\
g\left(\mu_i\right) &= \eta_{i} = A_{i} + \mathbf{X}_{i} \gamma + \sum_{j=1}^{J} f_{j}\left(x_{ij}\right)
\end{aligned}
\]

where \(g(\cdot)\) is the link function, \(A\) is an offset,
\(\mathbf{X}_{i}\) is the \(i\) th row of a parametric model matrix,
\(\gamma\) is a vector of parameters for the parametric terms, \(f_{j}\)
is a smooth function of covariate \(x_{j}\),
\(y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)\) denotes that the
observations \(y_{i}\) are distributed as some member of the exponential
family of distributions (e.g., Gaussian, Gamma, Beta, Poisson) with mean
\(\mu_{i}\) and scale parameter \(\phi\).

The smooth functions \(f_{j}\) are represented in the model via
penalised splines basis expansions of the covariates, that are a
weighted sum of basis functions:

@@

where \(\beta_{jk}\) is the weight (coefficient) associated with the
\(k\) th basis function \(b_{j k}()\) evaluated at the covariate value
\(x_{ij}\) for the \(j\) th smooth function \(f_{j}\)\ldots{}

@@

A Gaussian process (GP) is a stochastic process that defines the
distribution over a collection of random variables indexed by a
continuous variable, that is \(\{f(t): t \in \mathcal{T}\}\) for some
index set \(\mathcal{T}\) (\citeproc{ref-rasmussen2005}{Rasmussen \&
Williams, 2005};
\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
2023}).

Note that nonparametric does not mean there aren't parameters, it means
that there are infinitely many parameters.

From
\href{https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp}{brms
documentation}: A GP is a stochastic process, which describes the
relation between one or more predictors
\(x=\left(x_1, \ldots, x_d\right)\) and a response \(f(x)\), where \(d\)
is the number of predictors. A GP is the generalization of the
multivariate normal distribution to an infinite number of dimensions.
Thus, it can be interpreted as a prior over functions. The values of
\(f()\) at any finite set of locations are jointly multivariate normal,
with a covariance matrix defined by the covariance kernel
\(k_p\left(x_i, x_j\right)\), where \(p\) is the vector of parameters of
the GP:

\[
\left(f\left(x_1\right), \ldots f\left(x_n\right) \sim \operatorname{MVN}\left(0,\left(k_p\left(x_i, x_j\right)\right)_{i, j=1}^n\right)\right.
\]

The smoothness and general behaviour of the function \(f\) depends only
on the choice of covariance kernel, see
\url{https://mc-stan.org/docs/functions-reference/matrix_operations.html\#gaussian-process-covariance-functions}.
It ensures that values that are close together in the input space will
be mapped to similar output values\ldots{} From this perspective, \(f\)
is a realisation of an infinite dimensional normal distribution:

@@

where \(C\) is a covariance kernel with hyperparameters \(\lambda\) that
defines the covariance between two function values \(f\left(t_1\right)\)
and \(f\left(t_2\right)\) for two time points \(t_1\) and \(t_2\)
(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). Similar to
the different choices of the basis function for splines, different
choices of the covariance kernel lead to different GPs. In this article,
we consider the squared-exponential (a.k.a. Gaussian or radial basis
function) kernel, which computes the squared distance between points and
converts it into a measure of similarity. It is defined as:

\[
C(\lambda):=C\left(t_1, t_2, \sigma, \gamma\right):=\sigma^2 \exp \left(-\frac{\left(t_1-t_2\right)^2}{2 \gamma^2}\right)
\]

with hyperparameters \(\lambda=(\sigma, \gamma)\), expressing the
overall scale of GP and the length-scale, respectively
(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). The
advantages of this kernel are that it is computationally efficient and
(infinitely) smooth making it a reasonable choice for the purposes of
the present article. Here again, \(\lambda\) hyperparameters are
estimated from the data, along with all other model parameters.

@@

Following the approach used by Sassenhagen \& Draschkow
(\citeproc{ref-sassenhagen2019}{2019}) and Rousselet
(\citeproc{ref-rousselet_using_2025}{2025}), we simulated EEG data
stemming from two conditions, one with noise only, and the other with
noise + signal. As in previous studies, the noise was generated by
superimposing 50 sinusoids at different frequencies, following an
EEG-like spectrum (see details and code in
\citeproc{ref-yeung2004}{Yeung et al., 2004})\ldots{}

For simulating EEG data, see
\url{https://github.com/GRousselet/onsetsim} and
\url{https://github.com/unfoldtoolbox/UnfoldSim.jl}\ldots{}

Copy-pasted from Rousselet (\citeproc{ref-rousselet_using_2025}{2025}):
``The signal was a truncated Gaussian defining an objective onset at 160
ms, a maximum at 250 ms, and an offset at 342 ms. It was created by
getting the probability density function of a standard normal with 93
points from x= 1.5 to 1.5, rescaled in amplitude between 0 and 1 and
padded with 79 zeros to the left and to the right, to form a time series
of 251 points, from 0 to 500 ms, in steps of 2 ms (500 Hz sampling
rate).''

@@

Models were fitted using the \texttt{brms} package
(\citeproc{ref-burkner_brms_2017}{Bürkner, 2017},
\citeproc{ref-burkner_advanced_2018}{2018};
\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019})\ldots{}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting the GP}
\NormalTok{meta\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# using by{-}participant SD of ERPs across trials}
\NormalTok{    eeg }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{gp}\NormalTok{(x, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
\NormalTok{        (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ participant),}
    \AttributeTok{data =}\NormalTok{ summary\_df,}
    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{20}\NormalTok{),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/meta\_gp.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

@@

@@

Now, we want to test whether and when decoding performance is above
chance level (0.5 for a binary decoding task). This one is
computationally very intense\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting a GAM with a 2D smooth function}
\NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# 2D thin{-}plate spline (tp) to model smooth interactions between training and testing time}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{s}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
    \AttributeTok{data =}\NormalTok{ timegen\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/timegen\_gam.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

>>>>

\usepackage{mathtools}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

Here are some useful references to be discussed
(\citeproc{ref-combrisson_exceeding_2015}{Combrisson \& Jerbi, 2015};
\citeproc{ref-ehinger_unfold_2019}{Ehinger \& Dimigen, 2019};
\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021a},
\citeproc{ref-frossard2022}{2022}; \citeproc{ref-gramfort2013}{Gramfort,
2013}; \citeproc{ref-hayasaka_validating_2003}{Hayasaka, 2003};
\citeproc{ref-luck_how_2017}{Luck \& Gaspelin, 2017};
\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al., 2019};
\citeproc{ref-pernet2015}{Pernet et al., 2015};
\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al., 2023};
\citeproc{ref-rousselet_using_2025}{Rousselet, 2025})\ldots{} See also
(\citeproc{ref-maris2011}{Maris, 2011})\ldots{} and
(\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}) (history of
cluster-based approaches and using a data split?)\ldots{} Cluster
failure (\citeproc{ref-eklund2016}{Eklund et al., 2016})\ldots{}

@@

\[
\begin{aligned} 
y_{i} &\sim \mathrm{EF}\left(\mu_{i}, \phi\right)\\
g\left(\mu_i\right) &= A_{i} + \mathbf{X}_{i} \gamma + \sum_{j=1}^{J} f_{j}\left(x_{ij}\right)
\end{aligned}
\]

where \(y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)\) denotes that
the observations \(y_{i}\) are distributed as some member of the
exponential family of distributions (e.g., Gaussian, Gamma, Beta,
Poisson) with mean \(\mu_{i}\) and scale parameter \(\phi\);
\(g(\cdot)\) is the link function, \(A\) is an offset,
\(\mathbf{X}_{i}\) is the \(i\)th row of a parametric model matrix,
\(\gamma\) is a vector of parameters for the parametric terms, \(f_{j}\)
is a smooth function of covariate \(x_{j}\). The smooth functions
\(f_{j}\) are represented in the model via penalised splines basis
expansions of the covariates, that are a weighted sum of basis
functions:

@@

where \(\beta_{jk}\) is the weight (coefficient) associated with the
\(k\)th basis function \(b_{jk}()\) evaluated at the covariate value
\(x_{ij}\) for the \(j\)th smooth function \(f_{j}\). Splines'
coefficients are penalised\ldots{}

@@

A Gaussian process (GP) is a stochastic process that defines the
distribution over a collection of random variables indexed by a
continuous variable, that is \(\{f(t): t \in \mathcal{T}\}\) for some
index set \(\mathcal{T}\) (\citeproc{ref-rasmussen2005}{Rasmussen \&
Williams, 2005};
\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
2023}). Whereas Bayesian linear regression outputs a distribution over
the parameters of some predefind parametric model, the GP approach, in
contrast, is a non-parametric approach, in that it finds a distribution
over the possible functions that are consistent with the observed data.
However, note that nonparametric does not mean there aren't parameters,
it means that there are infinitely many parameters.

From
\href{https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp}{brms
documentation}: A GP is a stochastic process, which describes the
relation between one or more predictors
\(x=\left(x_{1}, \ldots, x_{d}\right)\) and a response \(f(x)\), where
\(d\) is the number of predictors. A GP is the generalization of the
multivariate normal distribution to an infinite number of dimensions.
Thus, it can be interpreted as a prior over functions. The values of
\(f()\) at any finite set of locations are jointly multivariate normal,
with a covariance matrix defined by the covariance kernel
\(k_p\left(x_i, x_j\right)\), where \(p\) is the vector of parameters of
the GP:

\[
\left(f\left(x_{1}\right), \ldots f\left(x_{n}\right) \sim \operatorname{MVN}\left(0, \left(k_p\left(x_{i}, x_{j}\right)\right)_{i, j=1}^{n}\right)\right.
\]

The smoothness and general behaviour of the function \(f\) depends only
on the choice of covariance kernel, which ensures that values that are
close together in the input space will be mapped to similar output
values\ldots{}

From this perspective, \(f\) is a realisation of an infinite dimensional
normal distribution:

@@

where \(C\) is a covariance kernel with hyperparameters \(\lambda\) that
defines the covariance between two function values \(f\left(t_1\right)\)
and \(f\left(t_2\right)\) for two time points \(t_1\) and \(t_2\)
(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). Similar to
the different choices of the basis function for splines, different
choices of the covariance kernel lead to different GPs. In this article,
we consider the squared-exponential (a.k.a. radial basis function)
kernel, which computes the squared distance between points and converts
it into a measure of similarity. It is defined as:

\[
C(\lambda) := C\left(t_1, t_2, \sigma, \gamma\right) := \sigma^2 \exp \left(-\frac{||t_1-t_2||^{2}}{2 \gamma^2}\right)
\]

with hyperparameters \(\lambda = (\sigma, \gamma)\), expressing the
overall scale of GP and the length-scale, respectively
(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). The
advantages of this kernel are that it is computationally efficient and
(infinitely) smooth making it a reasonable choice for the purposes of
the present article. Here again, \(\lambda\) hyperparameters are
estimated from the data, along with all other model parameters.

Taken from
\url{https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10c-Gaussian-processes.html}:
For a given vector \(\mathbf{x}\), we can use the kernel to construct
finite multi-variate normal distribution associated with it like so:

\[
\mathbf{x} \mapsto_{G P} \operatorname{MVNormal}(m(\mathbf{x}), k(\mathbf{x}, \mathbf{x}))
\]

where \(m\) is a function that specifies the mean for the distribution
associated with \(\mathbf{x}\). This mapping is essentially the Gaussian
process: a systematic association of vectors of arbitrary length with a
suitable multi-variate normal distribution.

@@

Following the approach used by Sassenhagen \& Draschkow
(\citeproc{ref-sassenhagen2019}{2019}) and Rousselet
(\citeproc{ref-rousselet_using_2025}{2025}), we simulated EEG data
stemming from two conditions, one with noise only, and the other with
noise + signal. As in previous studies, the noise was generated by
superimposing 50 sinusoids at different frequencies, following an
EEG-like spectrum (see details and code in
\citeproc{ref-yeung2004}{Yeung et al., 2004}). As in Rousselet
(\citeproc{ref-rousselet_using_2025}{2025}), the signal was generated
from truncated Gaussian with an objective onset at 160 ms, a peak at 250
ms, and an offset at 342 ms. We simulated this signal for 250 timesteps
between 0 and 0.5s, akin to \& 500 Hz sampling rate. We simulated such
data for a group of 20 participants with 50 trials per participant and
condition.

@@

Models were fitted using the \texttt{brms} package
(\citeproc{ref-burkner_brms_2017}{Bürkner, 2017a},
\citeproc{ref-burkner_advanced_2018}{2018a};
\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019})\ldots{}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting the GP}
\NormalTok{meta\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# using by{-}participant SD of ERPs across trials}
\NormalTok{    eeg }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{gp}\NormalTok{(x, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
\NormalTok{        (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ participant),}
    \AttributeTok{data =}\NormalTok{ summary\_df,}
    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{20}\NormalTok{),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/meta\_gp.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# loading the reticulate package}
\FunctionTok{library}\NormalTok{(reticulate)}

\CommentTok{\# importing the numpy and mne python modules}
\NormalTok{np }\OtherTok{\textless{}{-}} \FunctionTok{import}\NormalTok{(}\StringTok{"numpy"}\NormalTok{)}
\NormalTok{mne }\OtherTok{\textless{}{-}} \FunctionTok{import}\NormalTok{(}\StringTok{"mne"}\NormalTok{)}
\NormalTok{sklearn }\OtherTok{\textless{}{-}} \FunctionTok{import}\NormalTok{(}\StringTok{"sklearn"}\NormalTok{)}

\CommentTok{\# defining the function in R (it will be executed in Python)}
\NormalTok{mne\_decoding }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{ (X, labels, }\AttributeTok{ncores =} \DecValTok{1}\NormalTok{) \{}

    \CommentTok{\# converting R dataframe to NumPy array (reshaping if needed)}
    \CommentTok{\# X should be a matrix before conversion}
\NormalTok{    X\_np }\OtherTok{\textless{}{-}}\NormalTok{ np}\SpecialCharTok{$}\FunctionTok{array}\NormalTok{(X)}

    \ControlFlowTok{if}\NormalTok{ (}\FunctionTok{length}\NormalTok{(}\FunctionTok{dim}\NormalTok{(X\_np) ) }\SpecialCharTok{==} \DecValTok{2}\NormalTok{) \{}

        \CommentTok{\# adding a second dimension (channels) if missing}
\NormalTok{        X\_np }\OtherTok{\textless{}{-}}\NormalTok{ np}\SpecialCharTok{$}\FunctionTok{expand\_dims}\NormalTok{(X\_np, }\AttributeTok{axis =} \FunctionTok{as.integer}\NormalTok{(}\DecValTok{1}\NormalTok{) )}

\NormalTok{    \}}

    \CommentTok{\# defining the classifier}
\NormalTok{    clf }\OtherTok{\textless{}{-}}\NormalTok{ sklearn}\SpecialCharTok{$}\NormalTok{linear\_model}\SpecialCharTok{$}\FunctionTok{LogisticRegression}\NormalTok{(}\AttributeTok{solver =} \StringTok{"liblinear"}\NormalTok{)}

    \CommentTok{\# sliding the estimator on all time frames}
\NormalTok{    time\_decod }\OtherTok{\textless{}{-}}\NormalTok{ mne}\SpecialCharTok{$}\NormalTok{decoding}\SpecialCharTok{$}\FunctionTok{SlidingEstimator}\NormalTok{(}
\NormalTok{        clf, }\AttributeTok{n\_jobs =} \FunctionTok{as.integer}\NormalTok{(ncores),}
        \AttributeTok{scoring =} \StringTok{"roc\_auc"}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{TRUE}
\NormalTok{        )}

    \CommentTok{\# or using N{-}fold cross{-}validation}
\NormalTok{    scores }\OtherTok{\textless{}{-}}\NormalTok{ mne}\SpecialCharTok{$}\NormalTok{decoding}\SpecialCharTok{$}\FunctionTok{cross\_val\_multiscore}\NormalTok{(}
\NormalTok{        time\_decod,}
\NormalTok{        X\_np,}
\NormalTok{        labels,}
        \AttributeTok{cv =} \FunctionTok{as.integer}\NormalTok{(}\DecValTok{4}\NormalTok{),}
        \AttributeTok{n\_jobs =} \FunctionTok{as.integer}\NormalTok{(ncores),}
        \AttributeTok{verbose =} \ConstantTok{TRUE}
\NormalTok{        )}

    \CommentTok{\# returning the scores (averaged over CV folds)}
    \FunctionTok{return}\NormalTok{ (scores)}

\NormalTok{\}}

\CommentTok{\# listing all participants}
\NormalTok{participants }\OtherTok{\textless{}{-}} \FunctionTok{unique}\NormalTok{(raw\_df}\SpecialCharTok{$}\NormalTok{participant)}

\CommentTok{\# initialising empty decoding results}
\NormalTok{group\_decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{()}

\CommentTok{\# running decoding for each participant}
\ControlFlowTok{for}\NormalTok{ (ppt }\ControlFlowTok{in}\NormalTok{ participants) \{}

    \CommentTok{\# printing progress}
    \FunctionTok{print}\NormalTok{(ppt)}

    \CommentTok{\# retrieve data from one participant}
\NormalTok{    ppt\_data }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{filter}\NormalTok{(participant }\SpecialCharTok{==}\NormalTok{ ppt) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{participant) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ x, }\AttributeTok{values\_from =}\NormalTok{ y) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{condition, }\SpecialCharTok{{-}}\NormalTok{trial)}

    \CommentTok{\# extracting the labels}
\NormalTok{    labels }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{filter}\NormalTok{(participant }\SpecialCharTok{==}\NormalTok{ ppt) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{participant) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ x, }\AttributeTok{values\_from =}\NormalTok{ y) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{pull}\NormalTok{(condition) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{as.numeric}\NormalTok{()}

    \CommentTok{\# extracting the timesteps}
\NormalTok{    timesteps }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{filter}\NormalTok{(participant }\SpecialCharTok{==}\NormalTok{ ppt) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{participant) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{pull}\NormalTok{(x) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{unique}\NormalTok{()}

    \CommentTok{\# running the decoding}
\NormalTok{    decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
        \FunctionTok{mne\_decoding}\NormalTok{(}\AttributeTok{X =}\NormalTok{ ppt\_data, }\AttributeTok{labels =}\NormalTok{ labels}\DecValTok{{-}1}\NormalTok{)}
\NormalTok{        ) }\SpecialCharTok{\%\textgreater{}\%}
        \CommentTok{\# computing the average over CV folds}
        \FunctionTok{summarise}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{where}\NormalTok{(is.numeric), mean) )}

    \CommentTok{\# appending to previous results}
\NormalTok{    group\_decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(group\_decoding\_scores, decoding\_scores)}

\NormalTok{\}}

\CommentTok{\# saving the scores}
\FunctionTok{saveRDS}\NormalTok{(}\AttributeTok{object =}\NormalTok{ group\_decoding\_scores, }\AttributeTok{file =} \StringTok{"results/decoding\_scores.rds"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# importing the decoding scores}
\NormalTok{group\_decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\AttributeTok{file =} \StringTok{"results/decoding\_scores.rds"}\NormalTok{)}

\CommentTok{\# extracting the timesteps}
\NormalTok{timesteps }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
    \CommentTok{\# filter(participant == ppt) \%\textgreater{}\%}
    \CommentTok{\# select({-}participant) \%\textgreater{}\%}
    \FunctionTok{pull}\NormalTok{(x) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{unique}\NormalTok{()}

\CommentTok{\# plotting it}
\NormalTok{group\_decoding\_scores }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{t}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{data.frame}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}
        \AttributeTok{mean =} \FunctionTok{rowMeans}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{20}\NormalTok{) ),}
        \AttributeTok{se =} \FunctionTok{apply}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{20}\NormalTok{), }\DecValTok{1}\NormalTok{, }\ControlFlowTok{function}\NormalTok{ (x) }\FunctionTok{sd}\NormalTok{(x) }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(}\FunctionTok{length}\NormalTok{(x) ) )}
\NormalTok{        ) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lower =}\NormalTok{ mean }\SpecialCharTok{{-}}\NormalTok{ se, }\AttributeTok{upper =}\NormalTok{ mean }\SpecialCharTok{+}\NormalTok{ se) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{time =}\NormalTok{ timesteps) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time, }\AttributeTok{y =}\NormalTok{ mean) ) }\SpecialCharTok{+}
    \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{geom\_ribbon}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{ymin =}\NormalTok{ lower, }\AttributeTok{ymax =}\NormalTok{ upper), }\AttributeTok{fill =} \StringTok{"steelblue"}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{color =} \StringTok{"steelblue"}\NormalTok{, }\AttributeTok{size =} \FloatTok{0.8}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (s)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Decoding accuracy (ROC AUC)"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\caption{Exemplary group-level average timecourse of binary decoding
accuracy (ROC AUC). Decoding was performed using MNE-Python.}

{\centering \includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/unnamed-chunk-2-1.pdf}

@@

Now, we want to test whether and when decoding performance is above
chance level (0.5 for a binary decoding task). These two models are
computationally heavier to fit (more observations and 2D smooth
functions)\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting a GAM with two temporal dimensions}
\NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# 2D thin{-}plate spline (tp) to model smooth interactions between training and testing time}
    \CommentTok{\# auc \textasciitilde{} s(train\_time, test\_time, bs = "tp", k = 10),}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
    \AttributeTok{data =}\NormalTok{ timegen\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"}
\NormalTok{    )}
